Augmenting hierarchical time series through clustering: \\Is there an optimal way for forecasting?

### 2024-04-05 manuscript/manuscript.tex
minor
update

% \usepackage{natbib}
\usepackage{amsfonts,amsmath,bbm,bm,xcolor,booktabs,hyperref,amsthm}
\usepackage{geometry}
\usepackage{subfig}
\geometry{a4paper,scale=0.8}
\usepackage{setspace}
\setstretch{1.5}
\let\code=\texttt
\let\proglang=\textsf
\setcounter{MaxMatrixCols}{20}

@@

  \title{Constructing hierarchies}

@@

@@

  \begin{keyword}
  Forecasting \sep
  Hierarchical time series \sep

  \end{keyword}

@@

\section{Methodology}
\label{sec:method}

\subsection{Framework}

1. construct hierarchies based on historic observations
2. forecast and reconciliation
3. evaluation

\subsection{Constructing hierarchies by clustering}

\cite{aghabozorgiTimeseriesClusteringDecade2015a} summarised the four components of time series clustering, including time series representations, distance measures, clustering algorithms and evaluation. In our framework, the first three components are involved in the constructing hierarchies procedure. The last component includes evaluation of reconciled forecasts of the original hierarchy.

\subsubsection*{Time series representations}

Time-series representations represent raw time series in another space through techniques like feature extraction or dimensionality reduction. By utilising diverse time series representations, we are able to obtain distinct clusters which offer varied perspectives on the same dataset. 
Despite the absence of the theoretical evidence regarding how groups in hierarchical time series affect forecast reconciliation performance, we seek to explore multiple time series representations and discern potential correlations between hierarchical structure and the efficacy of forecast reconciliation.

Considering the impracticality of exploring every conceivable time series representation, we focus on four key ones: raw time series, in-sample error, features of time series, and features of in-sample error.
The inclusion of raw time series is motivated by its simplicity and broad applicability.
A crucial element contributing to the success of the minimum trace method proposed by \cite{wickramasuriyaOptimalForecastReconciliation2019} lies in estimating the covariance matrix of base forecast errors based on in-sample fit error. Hence, we consider in-sample error as a representation to examine how the structure within the error series influences reconciliation performance. Notably, to the best of our knowledge, we are the first to utilise in-sample error as a time series representation in the context of forecast reconciliation literature.
The time series and in-sample error representations are standardised on a per-series basis to eliminate the impact of scale variations.
Features, widely employed in capturing time series characteristics across literature, play a pivotal role in various time series applications, including clustering (\citealp{tianoFeatTSFeaturebasedTime2021}) and forecasting (\citealp{wangUncertaintyEstimationFeaturebased2022, liFeaturebasedIntermittentDemand2023}). In our exploration, we incorporate features of both time series and in-sample fit error as representations. These representations allow us to glean insights into the diverse aspects of hierarchical time series data and enhance our understanding of how different structures contribute to the forecast reconciliation performance.

\subsubsection*{Distance measures}

Distance measures serve as crucial tools for assessing the similarity between two series, forming the foundation for clustering algorithms to detect clusters within datasets. In the realm of time series clustering literature, we consider two widely applied distance measures: Euclidean distance and dynamic time warping (DTW) (\citealp{warrenliaoClusteringTimeSeries2005}).

Given the limited number of time series in comparison to the variable dimensions (i.e., the length of time series and the number of time series features) within hierarchical time series in our applications, dimension reduction on the aforementioned representations becomes imperative when employing Euclidean distance. Failure to undertake dimension reduction may result in undesirable clustering outcomes due to the challenges posed by the curse of dimensionality. To address this, we perform principal component analysis, extracting the first few principal components that collectively explain at least 80\% of the variance within the data.

In contrast, DTW (\citealp{sakoeDynamicProgrammingAlgorithm1978}), demonstrates reduced sensitivity to the curse of dimensionality. Unlike the one-to-one point comparison inherent in Euclidean distance, DTW accommodates time series of varying lengths through many-to-one comparisons. This adaptive approach allows for the recognition of time series with similar shapes, even in the presence of signal transformations such as shifting and/or scaling.

\subsubsection*{Clustering algorithms}
\label{sec:clustering}

Regarding clustering algorithms, we consider two common approaches, k-Medoids and agglomerative hierarchical clustering. K-Medoids algorithm is one of the classical partitioning clustering algorithms. The idea behind partitioning clustering is the minimisation of the total distance between all samples in a cluster from their cluster centres.  We consider k-Medoids instead of the well known k-Means because k-Means use mean vector of samples within a cluster as cluster centre, which is not trivial when using DTW distance measure. In contrast, k-Medoids select one of the samples in the cluster as cluster centre. We use the k-Meodoids variants partitioning around medoids (PAM, \citealp{PartitioningMedoidsProgram1990}). The most important parameter for k-Medoids algorithm is the number of cluster. We follow the suggestion of \cite{PartitioningMedoidsProgram1990} to choose the optimal number of cluster by average silhouette width (ASW), which is a popular cluster validation index to estimate the number of clusters. Given a clustering $\mathcal{C}$, the silhouette width for the $i$th sample is calculated as 
\[
  SW(i) = \frac{b(i)-a(i)}{\max\{a(i), b(i)\}},  
\]
where $a(i)$ is the average distance of $i$the sample to samples in the cluster to which it is assigned, and $b(i)$ is the average distance of $i$th sample to the samples in the nearest cluster it is not assigned. A large value of silhouette means that the corresponding sample is much closer to the samples to the samples in its own cluster than to the neighbouring one. Silhouette width is zero for the sample which is the only sample in its cluster. ASW is calculated as the average of silhouette widths of all samples. By iterating over all possible number of clusters, we choose the clustering result $\mathcal{C}$ that maximises the ASW as the optimal result. One drawback of this approach is that ASW does not exist when there is only one cluster, meaning that it can not determine whether there are well separated clusters or not. Gap statistics (\citealp{tibshiraniEstimatingNumberClusters2002}) can address this issue, but it can not be used with DTW.

hierarchical clustering

\subsection{Constructing hierarchies randomly}

pure random

random based on specific structure

\subsection{Forecast combination}

\section{Simulation}
\label{sec:simulation}

In this section, we demonstrate that hierarchies constructed by clustering may not outperform hierarchies constructed randomly even if there are true clusters in the bottom level.

>>>>

% \usepackage{natbib}
\usepackage{amsfonts,amsmath,bbm,bm,xcolor,booktabs,hyperref,amsthm}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{subfig}
\geometry{a4paper,scale=0.8}
\usepackage{setspace}
\setstretch{1.5}
\let\code=\texttt
\let\proglang=\textsf
\setcounter{MaxMatrixCols}{20}
\usepackage{enumitem}

% Define a new counter for your research questions
\newcounter{researchquestion}
\setcounter{researchquestion}{0}

% Define the environment for the research questions
\newenvironment{researchquestion}{%
    \refstepcounter{researchquestion}%
    \par\medskip\noindent%
    \textbf{RQ\theresearchquestion.}~%
}{\medskip}

@@

  %\title{On the performance of hierarchy construction: impacts of structure and clustering}
\title{Augmenting hierarchical time series through clustering: \\Is there an optimal way for forecasting?}

  \author[label1]{Bohan Zhang\corref{cor1}}
  \address[label1]{School of Economics and Management, Beihang University, Beijing, China}
  \ead{zhangbohan@buaa.edu.cn}
  \cortext[cor1]{Corresponding author.}
  \author[label2]{Anastasios Panagiotelis}
  \address[label2]{The University of Sydney Business School, NSW 2006, Australia}
  \author[label3]{Han Li}
  \address[label3]{Department of Economics, The University of Melbourne, VIC 3010, Australia}

@@

    Forecast reconciliation has attracted significant research interest in recent years, with most studies relying on pre-defined hierarchies constructed with time series metadata. With the goal of improving forecast accuracy in mind, we extend and contribute to the emerging research on the clustering-based reconciliation method by proposing a novel framework for hierarchy construction. This framework offers three approaches: cluster hierarchies, random hierarchies, and combination hierarchies. Utilizing the proposed approaches, we investigate the individual contributions of two primary factors, namely ``grouping'' and ``structure'', to the performance of forecast reconciliation.  Through a simulation study and experiments on two real-world datasets, we demonstrate the practical efficacy of different hierarchy construction approaches. Our findings provide new insights into the dynamics between ``grouping'' and ``structure'', which lead to an improved understanding of forecast reconciliation.\\

@@

  \begin{keyword}
  Forecast reconciliation \sep
  Hierarchical time series \sep
  Clustering \sep
  Hierarchy construction \sep
  Forecast combination
  \end{keyword}

@@

Applications where some variables are aggregates of one another, or so-called \textit{hierarchical time series (HTS)}, are found in many forecasting problems ranging from supply chain management (\citealp{syntetosSupplyChainForecasting2016}) to tourism planning (\citealp{kourentzesCrosstemporalCoherentForecasts2019}), electrical load forecasting (\citealp{jeonProbabilisticForecastReconciliation2019}), and retail demand forecasting (\citealp{makridakisM5AccuracyCompetition2022}). In recent decades, there has been an increasing interest in hierarchical forecasting, primarily driven by the success of the optimal reconciliation framework (\citealp{hyndmanOptimalCombinationForecasts2011,wickramasuriyaOptimalForecastReconciliation2019, panagiotelisProbabilisticForecastReconciliation2023}). The original motivation for forecast reconciliation was to ensure forecasts are \textit{coherent}, that is they respect the aggregation constraints implied by the hierarchical structure. Coherent forecasts faciliate aligned decisions by agents acting upon different variables within the hierarchy. For example, consider a retail setting, where a warehouse manager supplies stock to individual store managers within their region. Forecasts could be incoherent when the warehouse manager forecasts low total demand while store managers forecast high demand, leading to supply shortages. Numerous case studies in the literature demonstrate that reconciliation approaches not only yield coherent forecasts but also enhance overall forecast performance (\citealp{AthanasopoulosForecastReconciliationReview2023}).

%The primary objective of hierarchical forecasting is to produce coherent forecasts for the series within a given HTS, which we refer to as \textit{original hierarchy}, ensuring that decision-makers at different levels of the hierarchy can analyze and make informed decisions based on aligned expectations for the future.
A limitation in the overwhelming majority of the forecast reconciliation studies is that the structure of the hierarchy is taken as \textit{given}. This structure usually includes \textit{bottom level series}, an overall aggregate or \textit{top level series}, with various aggregation schemes used to construct \textit{middle levels series}. Typically, middle levels are formed according to inherent attributes of the bottom-level series, such as geographical location, gender, product category, travel purpose, and others. We refer to this type of structure as the \textit{natural hierarchy}. While in some forecasting applications, decisions must be made with respect to the natural hierarchy, in other settings there might be some flexibility in determining how bottom levels are aggregated into middle levels. It should be noted that,  very little attention has been paid to whether middle level series can be constructed in a way that leads to further improvements in forecast accuracy relative to a given natural hierarchy. To the best of our knowledge, only \cite{pangHierarchicalElectricityTime2018}, \cite{liForecastReconciliationApproach2019}, \cite{pangHierarchicalElectricityTime2022}, and \cite{matteraImprovingOutofSampleForecasts2023} have attempted to \textit{construct} middle level series in a data-driven way which ultimately improved forecast accuracy. All three of these works use time series clustering to construct hierarchies in a manner that is somewhat ad hoc. Our work conducts a more thorough investigation of issues faced when  constructing hierarchical structures in forecasting reconciliation. In particular, we address the following four research questions:%via time series clustering. %We also assess and disentangle the impact of

%This limitation is evident in the absence of any discussion regarding cluster-based approaches in the recent comprehensive review of forecast reconciliation by \cite{AthanasopoulosForecastReconciliationReview2023}. However, the following questions arise when pursuing superior forecast performance, motivating the investigation presented in this paper.

\begin{researchquestion}
In terms of forecast performance, is the natural hierarchy better than a two-level hierarchy (consisting of only top and bottom time series)? 
%In terms of forecast performance, does the natural hierarchy outperform a two-level hierarchy of only top and bottom time series?
\end{researchquestion}

%Two observations have been consistently demonstrated in empirical studies on forecast reconciliation. First, the accuracy of base forecasts, which are influenced by the individual time series characteristics within the hierarchy, critically affects the overall performance of reconciliation. For instance, \cite{panagiotelisForecastReconciliationGeometric2021} show that greater performance improvement can be achieved when bias correction for base forecasts is involved. 
%Under the reconciliation framework, improvements in accuracy for some series often come at the expense of others (see \citealp{pritulargaStochasticCoherencyForecast2021}).
To investigate this question, we consider two widely used empirical HTS datasets; the first is Australian tourism demand, the second, cause-of-death mortality data.  Throughout the paper, all evaluations are carried out using the series common to all hierarchies; namely the top and bottom level series. In both datasets, we find that the natural hierarchy outperforms the two-level hierarchy, leading to our next research question:
\begin{researchquestion}
    If the use of middle level series in the ``natural'' hierarchy can lead to improvement in forecast accuracy, is it possible to construct hierarchies in a data-driven way that leads to further improvements in forecast accuracy?
\end{researchquestion}

The rationale behind the data-driven approach lies in grouping time series with similar patterns together, thereby creating middle-level series with enhanced signals and consequently, improved forecastability. Such arguments have been put forward by  \cite{pangHierarchicalElectricityTime2018}, \cite{liForecastReconciliationApproach2019}, \cite{pangHierarchicalElectricityTime2022}, and \cite{matteraImprovingOutofSampleForecasts2023}. However, these studies for the most part focus on a small number of (in some cases, a single) clustering techniques. In this paper, we take a more systematic approach by clustering time series using different representations (the original time series, forecast errors, features of both), different distance metrics (Euclidean, dynamic time warping), and different clustering paradigms ($k$-medioids, hierarchical). Using both empirical datasets, as well as a simulation study, we find evidence that constructing hierarchies via clustering can lead to improved forecasting performance, although the optimal clustering method depends on the dataset as well as the base forecasting and reconciliation method.

While the idea behind time series clustering is intuitively appealing, the increased accuracy when using clustering-based methods may be attributed to two factors. The first, which we refer to as     ``grouping'' is the idea that some correct subsets of series are chosen to form new middle-level series. This is the argument commonly made to support clustering-based hierarchy construction \citep[see e.g.][]{liForecastReconciliationApproach2019, pangHierarchicalElectricityTime2022, matteraImprovingOutofSampleForecasts2023}.  
%The idea that often motivates time series clustering is these ``groupings'' be composed of similar series. 
The second factor, which we refer to as the ``structure'' of the hierarchy, includes the number of middle level series, the depth of the hierarchy, and the distribution of group sizes in the middle layer(s). Evidence showing that clustering within a forecast reconciliation framework leads to improved forecast accuracy, does not disentangle contributions from these two factors. Indeed, clustering methods may only work in so far as they generate a larger number of base forecasts. This argument would be consistent with the interpretation of reconciliation as a forecast combination of ``direct'' and ``indirect'' forecasts (\citealp{hollymanUnderstandingForecastReconciliation2021}), since more middle-level series implies a greater number of indirect forecasts in the combination. This leads to our third research question:

\begin{researchquestion}
    Can the improved accuracy of cluster-based methods be attributed to grouping together similar time series, or to the structure of the hierarchy including the number of middle level series?
\end{researchquestion}

To investigate this question, we devise the following approach. We take a hierarchy found using a certain clustering method (or even the natural hierarchy), and then randomly permute the bottom level series (\textit{i.e.}, the leaf nodes of the hierarchical tree). Multiple new ``twin'' hierarchies are formed with an identical structure to the original hierarchy, but with permuted leaves. In this way, we keep the \textit{hierarchical structure} fixed, but alter how series are combined. This method can be thought of as an informal ``permutation type'' test \citep{welch1990}. Our main finding is that hierarchies constructed using clustering methods do not significantly outperform their random ``twins'', leading to the conclusion that the driver of forecast improvement is the enlarged number of series in the hierarchy and/or its structure, rather than similarities between the time series.

%Despite the intuitive appeal of clustering analysis, it is not straightforward to conclude that clustering itself is the primary contributor to the performance improvement illustrated in the literature. Except the factor of potentially accurate base forecasts of new series introduced by clustering, another factor  is the ``enriched structure'', meaning that a larger set of series is optimally combined to obtain the reconciled forecasts, leading to reduced model uncertainty and data uncertainty from forecast combination perspective (\citealp{hollymanUnderstandingForecastReconciliation2021}). The following question also motivates our research:
%\begin{researchquestion}
%Which factor, enriched structure or clustering, is the primary contributor to the performance improvement illustrated in clustering-based reconciliation approaches?
%\end{researchquestion}

%To address this question, we propose a random hierarchy construction approach. By randomly constructing ``twin'' hierarchies with the same tree structure but different permutations of bottom-level series as a contrast to a clustering-based hierarchy, we are able to isolate the influence of clustering algorithms. This allows us to inspect the individual contributions of the enriched structure and clustering technique to forecast reconciliation performance.

Finally, from a practical perspective, we investigate the role of forecast combination in cluster-based hierarchical forecasting. With multiple hierarchies available and inspired by the forecast combination literature (\citealp{wangForecastCombinations50year2022}), we consider the last research question

\begin{researchquestion}
    Does an equally-weighted combination of reconciled forecasts derived from multiple random hierarchies improve forecast reconciliation performance?
\end{researchquestion}

Note that forecast combination here differs from that of \cite{hollymanUnderstandingForecastReconciliation2021}, in that our approach averages not only different coherent forecasts, but also over hierarchies with completely different middle level series. This is possible since only coherent bottom and top level forecasts are averaged and evaluated.
\clearpage
In summary, this paper presents four main contributions:

\begin{itemize}
  \item We introduce a novel hierarchical forecast reconciliation framework centered on hierarchy construction.  Within this framework, we introduce and compare three distinct approaches: cluster-based hierarchies, hierarchies based on random permutation, and  forecast combinations across different hierarchies.
  \item In contrast to existing literature that often focuses on a single clustering technique, our study systematically investigates the effectiveness of various time series clustering implementations. This investigation involves the incorporation of four time series representations, two distance measures, and two clustering algorithms.
  \item We conduct experiments using two empirical datasets - the Australian tourism dataset and the U.S. cause-of-death mortality dataset as well as a synthetic dataset. The results allow for a comparison of different approaches to constructing hierarchies.
  \item By constructing random hierarchies through permutation of leaf nodes, we show that the hierarchical structure is the primary contributor to improvements in forecast reconciliation performance, rather than the grouping of similar bottom level series.

\end{itemize}

The rest of the paper is organized as follows. Section~\ref{sec:setup} describes the experimental setup including the datasets used, the reconciliation methods employed, the clustering techniques considered, and the rolling window evaluation with associated forecast performance metrics. Section~\ref{sec:natural} investigates RQ1, in particular the performance of the natural hierarchy compared to the two-level hierarchy. The permutation approach used throughout the paper is also introduced at this juncture. Section~\ref{sec:clustering} evaluate the performance of different clustering approaches in answering RQ2 and RQ3 via two empirical studies. To avoid the concern that clusters found in the empirical datasets are spurious, a simulation study is considered in Section~\ref{sec:simulation}. Section~\ref{sec:combination} covers the forecast combination approaches raised in RQ4. Section~\ref{sec:conclusion} concludes this paper with discussions on the findings and outlines future research directions.

%Section~\ref{sec:related_work} reviews related work in time series clustering, forecast reconciliation and clustering-based forecast reconciliation. In Section~\ref{sec:method} we introduce the proposed framework, elaborating on three approaches: cluster hierarchies, random hierarchies and combination hierarchies. The simulation study is presented in Section~\ref{sec:simulation}, while Section~\ref{sec:emp} details empirical experiments conducted on the Australian tourism dataset and U.S. mortality dataset. Finally, Section~\ref{sec:conclusion} concludes this paper with discussions on the findings and outlines future research directions.

\section{Experiment setup}\label{sec:setup}

\subsection{Dataset description}

We conduct our experiments on two empirical datasets throughout this paper. The first one is the monthly Australian domestic tourism dataset, covering the period from January 1998 to December 2016. The data is recorded as ``visitor nights'', representing the total number of nights spent by Australians away from home. In this dataset, the total visitor nights of Australia is geographically disaggregated into seven states and territories, which are further divided into $27$ zones, and then into $76$ regions. Additionally, each regional-level series is divided by four travel purposes. Overall, this dataset comprises a total of $555$ time series with $304$ of those at the bottom level.
Section 4 of \cite{wickramasuriyaOptimalForecastReconciliation2019} provided an in-depth explanation of this dataset.

The second dataset focuses on cause-of-death mortality in the U.S. We obtain monthly cause-specific death count data from the Center for Disease Control and Prevention (CDC) for the period between January 1999 and December 2019. The dataset, organized based on the 10th revision of the International Classification of Diseases (ICD) 113 Cause List\footnote{For more detailed information on the dataset, please refer to \url{https://wonder.cdc.gov/ucd-icd10-expanded.html}.}, forms a hierarchy containing $120$ time series, with $98$ of those being bottom-level series\footnote{To address the data suppression issue, we combined certain ICD codes to ensure all death counts are no less than 10. }. The top-level series represents the aggregated deaths from all causes, while the middle-level series are constructed based on major cause-of-death groups. As an example, \textit{Diseases of heart} (ICD code: I00--I09, I11, I13, I20--I51; 113 Cause List: GR113-054) is a middle-level series in the hierarchy, which contains bottom-level series \textit{Hypertensive heart disease} (I11; GR113-056) and \textit{Heart failure} (I50; GR113-067), among other circulatory diseases.  % To address issues with suppressed data values, we combine causes that contain suppressed values and share a parent cause. The death counts for these combined categories are calculated by subtracting the death counts of sibling causes from the death counts of their parent causes. This processing results in a refined dataset that includes $120$ time series with $98$ at the bottom level.

Figures~\ref{fig:tourism} and~\ref{fig:mortality} illustrate the top-level series and four bottom-level series for the tourism and mortality datasets, respectively. 
In the case of tourism data, the first three letters indicate geographical regions, and the last three denote travel purposes. For example, ``\textit{BEEHol}'' represents the visitor nights spent for holiday in the `Spa Country' region. As mentioned before, the death series are coded based on the ICD 113 Cause List. From the figures, we can see that the mortality dataset generally exhibits strong seasonality and trend, whereas the tourism dataset displays greater volatility and less obvious seasonal and trend components. We can also see that the patterns observed at the bottom level can be very different from those at the top and middle levels.

%Names of mortality series are ICD codes representing causes of deaths. For example, \code{GR113-085} represents deaths because of asthma.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/tourism.pdf}
    \caption{Visualization of selected time series from the tourism dataset. ``AEB'', ``BAB'', ``BEB'', and ``BEE'' represent the regions `New England North West`, `Peninsula', `Western Grampians` and ``Spa Country'', respectively. ``Bus'', ``Oth'', ``Vis'' and ``Hol'' denote travel purposes ``Business'', ``Other'', ``Visit'' and ``Holiday'', respectively.}
    \label{fig:tourism}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/mortality.pdf}
    \caption{{Visualization of selected time series from the death count dataset.} ``GR113-051'',  ``GR113-063'', ``GR113-085'', and ``GR113-091'' denote ``Parkinson disease'', ``All other forms of chronic ischemic heart disease'', ``Asthma'', and ``Diseases of appendix'', respectively.}
    \label{fig:mortality}
\end{figure}

\subsection{Trace minimization forecast reconciliation approach}

In this subsection, we describe the process of generating coherent forecasts using the trace minimization (MinT) reconciliation approach \citep{wickramasuriyaOptimalForecastReconciliation2019}, which will be employed throughout this paper. Consider a given hierarchy with $n$ time series and $m$ of them being in the bottom level.  Let vectors $\boldsymbol{b}_t$, $\boldsymbol{a}_t$, and $\boldsymbol{y}_t$ represent observations of the bottom level, the aggregated levels (\textit{i.e.} middle and top levels), and the entire hierarchy at time $t$, respectively.
They are linked through an $n\times m$ summing matrix $\boldsymbol{S}$, which can be decomposed into an identity matrix $\boldsymbol{I}_m$ and a constraint matrix $\boldsymbol{A}$ consisting of $0$ and $1$, such that, 
\[
  \boldsymbol{y}_t = \boldsymbol{S}\boldsymbol{b}_t = \begin{bmatrix}
    \boldsymbol{A} \\ \boldsymbol{I}_m 
  \end{bmatrix}  \boldsymbol{b}_t = \begin{bmatrix}
      \boldsymbol{a}_t \\ \boldsymbol{b}_t
  \end{bmatrix},
\]
where $\boldsymbol{A}$ represents the mapping from bottom-level time series to aggregated-level time series.

Assume our objective is to make $h$-step-ahead forecasts based on $T$ historic observations, we would first produce $h$-step-ahead unreconciled (or ``base'') forecasts $\hat{\boldsymbol{y}}_{T+h}$. Note that there are many different ways to compute such forecasts. In this paper, we adopt the Exponential Smoothing (ETS) method, a well-known univariate forecasting model widely used in both academia and industry \citep{ForecastingExponentialSmoothing}. 
The \code{forecast} (\citealp{forecast}) package in {R} (\citealp{R}) is implemented to produce base forecasts for each time series in the hierarchy.
Subsequently, the MinT reconciliation method is applied to produce coherent forecasts:
\[
    \tilde{\boldsymbol{y}}_{T+h} = (\boldsymbol{S}'\boldsymbol{W}_h^{-1}\boldsymbol{S})^{-1}\boldsymbol{S}'\boldsymbol{W}_h^{-1}\hat{\boldsymbol{y}}_{T+h},
\]
where $\boldsymbol{W}_h$ is the covariance matrix of $h$-step-ahead forecast errors, which is estimated using the MinT shrinkage estimator based on the in-sample one-step-ahead forecast errors (\citealp{wickramasuriyaOptimalForecastReconciliation2019}).

%\subsection{Grouping-based forecast reconciliation}
\subsection{Augmenting hierarchical time series through clustering}

%DELETE

%Instead of a We use the term ``augmentation'' to denote the process of creating middle-level series by aggregating subsets of bottom-level series. leading to a modified hierarchical time series $\boldsymbol{y}^c_t$, which can be represented as:
%leading to a modified hierarchical time series $\boldsymbol{y}^c_t$, which can be represented as:
%\[
%   \boldsymbol{y}^c_t = \boldsymbol{S}^c\boldsymbol{b}_t = \begin{bmatrix}
%        \boldsymbol{A}' &
%        \boldsymbol{C}_1 &
%        \cdots &
%        \boldsymbol{C}_K &
%        \boldsymbol{I}_m
    %\end{bmatrix}'\boldsymbol{b}_t,
%\] 
%where each $\boldsymbol{C}_k$ represents a newly created middle-level series, formed by aggregating a subset of bottom-level series $\boldsymbol{b}_t$. The vector $\boldsymbol{C}_k$ is of $m$-dimensional and its elements $c_{k, i} \in \{0, 1\}, i=1,\dots m$ indicate whether the $i$-th bottom-level series belongs to the aggregated subset. Notably, different augmentation approaches can result in different hierarchies with varying numbers of new middle-level series $K$.

%After new hierarchies are constructed, we produce $h$-step-ahead base forecasts based on $T$ observations for each time series in the new hierarchy, and apply forecast reconciliation with MinT shrinkage estimator to obtain reconciled forecasts $\tilde{\boldsymbol{y}}^c_{T+h}$. Given our specific interest in the performance of the two-level hierarchy, we extract reconciled forecasts of the two-level hierarchy from $\tilde{\boldsymbol{y}}_{T+h}^c$, \textit{i.e.},
%\[
%  \tilde{\boldsymbol{y}}_{T+h} = \tilde{\boldsymbol{y}}^c_{T+h, 1:(n-m) \cup (n + K - m + 1): (n+K)}.
%\]

%\subsection{Cluster hierarchy construction approach}

To the best of our knowledge, the following studies are the exclusive examples in the literature that propose to integrate forecast reconciliation with hierarchy construction via time series clustering.
\cite{pangHierarchicalElectricityTime2018} detect consumption patterns of electricity smart meter data based on X-means clustering algorithm, while \cite{pangHierarchicalElectricityTime2022} propose a multiple alternative clustering method to group similar electricity and solar power time series. 
\cite{liForecastReconciliationApproach2019} apply agglomerative clustering to cause-of-death time series, and
\cite{matteraImprovingOutofSampleForecasts2023} utilize Partition Around Medoids algorithms to unveil underlying structures in stock price indexes. 
These studies demonstrate superior forecast performance of the clustering-based hierarchies compared to the natural hierarchy or the two-level hierarchy.
However, these studies are limited in scope as they focus on a small number of  clustering techniques. 
%construct hierarchies through clustering algorithms and

Inspired by the comprehensive overview of time series clustering by \cite{aghabozorgiTimeseriesClusteringDecade2015a}, we consider various approaches based on three key components, namely time series representations, distance measures, and clustering algorithms. Our framework for creating a clustering approach is as follows: we first choose a time series representation, which can be the raw time series, the in-sample forecast error, features of the time series, or features of the in-sample forecast error. We then select a distance measure, either Euclidean distance or dynamic time warping. Once the representation and the distance measure are determined, we choose a clustering algorithm, either $k$-medoids or agglomerative hierarchical clustering, to form clusters in the dataset. In total, we consider 12 different clustering approaches in our experiments.
%, and evaluation measures.

% Time series clustering is an essential technique that finds applications across various domains, including finance (\citealp{DIAS2015852}), energy (\citealp{pangHierarchicalElectricityTime2018}), robotics (\citealp{gopalapillaiExperimentationAnalysisTime2014}), and others. It is commonly used to identify similar patterns within time series datasets, aiding in applications such as anomaly detection, pattern discovery, and recommendation.

\subsubsection{Time series representations}

The first step of time series clustering is to find appropriate time series representations. Time series representations transform time series data into another space through techniques like feature extraction or dimension reduction. By utilizing diverse time series representations, we are able to obtain distinct clusters which offer varied perspectives on the same dataset.
Considering the impracticality of exploring every conceivable time series representation, we focus on four key ones: raw time series without any transformation (hereafter referred to as ``time series''), in-sample one-step-ahead forecast error, features of time series, and features of in-sample one-step-ahead forecast error.
The inclusion of raw time series is motivated by its simplicity and broad applicability.
A crucial element contributing to the success of the MinT method lies in estimating the covariance matrix of base forecast errors based on in-sample one-step-ahead forecast error. Hence, we consider in-sample one-step-ahead forecast error as a representation to examine how the structure within the error series influences reconciliation performance. 
Raw time series and in-sample error representations are standardized to eliminate the impact of scale variations. This is done by normalizing each series by subtracting its mean and then dividing by the standard error.

Features, widely employed in capturing time series characteristics across literature, play a pivotal role in various time series applications, including clustering (\citealp{tianoFeatTSFeaturebasedTime2021}) and forecasting (\citealp{wangUncertaintyEstimationFeaturebased2022, liFeaturebasedIntermittentDemand2023}). In our exploration, we incorporate features of both raw time series and in-sample forecast error as representations. The included time series features, calculated by the \code{tsfeatures} (\citealp{tsfeatures}) in R, have been applied in several feature-based forecasting studies, \textit{e.g.}, \cite{montero-mansoFFORMAFeaturebasedForecast2020}.
Notably, to the best of our knowledge, we are the first to utilize in-sample forecast error and time series features as representations in the context of forecast reconciliation literature.
These representations allow us to gain insights into the diverse aspects of hierarchical time series data.

\subsubsection{Distance measures}

Distance measures serve as crucial tools for assessing the similarity/dissimilarity between two series, significantly influencing the clustering results.
We consider two widely applied distance measures: Euclidean distance and dynamic time warping (DTW). Due to the limited number of time series in comparison to the high dimensionality of individual time series (\textit{i.e.}, the length of time series and the number of time series features) in our applications, dimension reduction on the aforementioned representations becomes imperative when employing Euclidean distance. Failure to undertake dimension reduction may result in undesirable clustering outcomes due to the curse of dimensionality. To address this, we perform Principal Component Analysis (PCA), extracting the first few principal components that collectively explain at least 80\% of the variance within the representations.

In contrast, DTW (\citealp{sakoeDynamicProgrammingAlgorithm1978}), exhibits reduced sensitivity to the curse of dimensionality. Unlike Euclidean distance, which performs one-to-one point comparisons, DTW accommodates time series of varying lengths through many-to-one comparisons. This flexible approach allows for the recognition of time series with similar shapes, even in the presence of signal transformations such as shifting and/or scaling.

\subsubsection{Clustering algorithms}
Clustering algorithms, such as partitioning, hierarchical, grid-based, model-based clustering, among others, identify clusters by optimizing specific objective functions (\citealp{aghabozorgiTimeseriesClusteringDecade2015a}). These functions are calculated based on chosen time series representations and distance measures.
In this paper, we focus on two prevalent approaches, $k$-medoids and agglomerative hierarchical clustering. The $k$-medoids algorithm, a classical partitioning clustering method, aims to minimize the total distance between all samples within a cluster and their respective cluster centers. 
Unlike $k$-means which employs the mean vector of samples as the cluster center, $k$-medoids selects one sample within the cluster as the center. 
Specifically, we adopt the $k$-medoids variant, partitioning around medoids (PAM, \citealp{PartitioningMedoidsProgram1990}).
Following the recommendation of \cite{PartitioningMedoidsProgram1990}, we determine the optimal number of cluster using the average silhouette width (ASW), a popular cluster validation index.
ASW assesses the quality of clustering results by measuring the proximity of samples within a cluster compared to neighboring clusters.
Given a clustering result, the silhouette width for the $i$th sample is calculated as 
\[
  SW(i) = \frac{b(i)-a(i)}{\max\{a(i), b(i)\}},  
\]
where $a(i)$ represents the average distance of the $i$th sample to others in the same cluster, and $b(i)$ is the average distance to samples in the nearest cluster it is not assigned to. A higher silhouette width indicates greater proximity to samples within the same cluster than to those in neighboring clusters. The ASW is the average of silhouette widths across all samples. We select the clustering result that maximizes the ASW by iterating over all possible numbers of clusters. However, ASW has the limitation of being undefined when there is only one cluster.

On the other hand, agglomerative hierarchical clustering begins by considering each sample as a cluster, and then gradually merges clusters until all samples forms a single cluster. Recall that $m$ represents the number of bottom level series. 
This process results in a binary hierarchical tree with ($2m-1$) nodes. We employ Ward's linkage (\citealp{murtaghWardHierarchicalAgglomerative2014a}) to merge clusters, minimizing the increase of within-cluster variances at each step.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.45\textwidth]{figures/pamcluster.pdf}
    \hspace{1cm}
    \includegraphics[width=0.45\textwidth]{figures/aggcluster.pdf}
    \caption{\label{fig:cluster_example}Example clustering results of two clustering algorithms. Left panel displays example for $k$-medoids algorithm, and right panel displays example for agglomerative hierarchical clustering algorithm.}
\end{figure}
Illustrated in Figure~\ref{fig:cluster_example} are two example hierarchies generated by $k$-medoids and agglomerative clustering algorithms for an original hierarchy with four bottom-level series and one total series. These examples highlight the distinct behaviors of the two algorithms.
Firstly, $k$-medoids constructs a simple hierarchy with a single middle level, while hierarchical clustering generates multiple nested middle levels. 
Secondly, $k$-medoids produces a hierarchy with the same number of series as the optimal number of clusters, whereas hierarchical clustering yields multiple middle levels with ($m-2$) series.
As the number of bottom-level series increases, the structural differences become increasingly significant, and we aim to investigate how these differences influence the performance of forecast reconciliation.
